---
title: "MATH 216 Homework 4"
author: Nina Sonneborn
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Suggested packages; feel free to use more!
library(tidyverse)
library(USAboundaries)
library(maptools)
library(sp)
library(broom)
```



## Admistrative:

Please indicate

* Who you collaborated with:
* Roughly how much time you spent on this HW so far:
* The URL of the RPubs published URL [here](http://rpubs.com/nsonneborn/hw-4).
* What gave you the most trouble:
* Any comments you have:



## Question 1:

**Question**:  DHL, a German shipping company, has a ton of money to build a
shipping hub in the United States. Where should they build it?

**Deliverable**: A choropleth map of the lower 48 states (no Puerto Rico) where
the color corresponds to the distance of the given area to the shipping hub,
which is marked on the map with a large red dot.

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Perform your computations here:
# distance = sqrt(change_long_squared + change_in_lat_squared)

#Load the data
counties_sp <- us_counties()
counties_data <- counties_sp@data %>% mutate(n=1:3220)
centroids <- coordinates(counties_sp)
county_poly <- tidy(counties_sp, region = "geoid")

centroids <- as.data.frame(centroids) %>% 
  mutate(long = V1, lat = V2, n=1:3220) %>% 
  select(n, long, lat)

# for population data...
census <- read_csv('census_2010.csv') %>% 
  select(Geo_FIPS, SE_T001_001) %>% 
  mutate(population=SE_T001_001)

counties_data <- left_join(counties_data, centroids, by = "n")

counties <- left_join(county_poly, counties_data, by=c("id"="geoid")) %>%
  filter(!is.na(state_name)) %>%
  filter( !state_name %in% c("Alaska", "Hawaii", "Puerto Rico"))

counties1 <- left_join(county_poly, counties_data, by=c("id"="geoid")) %>% 
  filter(!state_name %in% c("Alaska", "Hawaii", "Puerto Rico"))

hub_data <- left_join(counties, census, by=c("id"="Geo_FIPS")) %>% 
  select(long.y, lat.y, population, id)

hub_data <- hub_data %>% mutate(long=long.y, lat = lat.y) %>% select(long, lat, population, id)

hub_data <-  hub_data %>% unique() %>% mutate(n=1:3108)
# Calculated weighted center twice: once for latitude, once for longitude.

find_hub <- hub_data %>%
  group_by(n) %>%
  filter(!is.na(population)) %>% #ignore a single missing value in row 2891
  mutate(weighted_lat = as.numeric(population)*lat) %>%
  mutate(weighted_long = as.numeric(population)*long)

find_hub %>% ungroup() %>% summarise(hub_lat = sum(weighted_lat)/sum(as.numeric(population)))

find_hub %>% ungroup() %>% summarise(hub_long = sum(weighted_long)/sum(as.numeric(population)))

hub_lat <- 37.5508
hub_long <- -91.649


# Find the population center
hub_data <- hub_data %>% 
  mutate(distance = sqrt((long - hub_long)^2 + (lat - hub_lat)^2))

hub_distance_data <-left_join(counties1, hub_data, by="id")

```

The formula for one-dimensional center of mass is [here](http://hyperphysics.phy-astr.gsu.edu/hbase/imgmec/cm.gif)
In the case of this 

```{r, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5}
# Here is a template of a ggplot call; feel free to change this to suit your
# needs. Once your computations above are complete, set the eval=TRUE on the
# line above to tell R Markdown to run this code block.
ggplot(data=hub_distance_data, aes(x=long.x, y=lat.x, group = group, fill=distance)) +
  # Plot choropleth polygons, low -> high being colored white -> dark green
  geom_polygon() +
  scale_fill_gradient(low="white", high="darkgreen") + 
  # Trace outlines of areas
  geom_path(size=0.1) +
  # Mark hub
  annotate("point", x=hub_long, y=hub_lat, col="red", size=5) +
  # Correct aspect ratio for map
  coord_map() +
  labs(title = "Distance from Shipping Hub")
```





